---
title: "CLT_WUP"
author: "Kimberly Aldana"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

kkardashtemp  see code below

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(lattice)
```

```{r}
Homers_data <- read.csv(file="Homers_data.csv")

my1000 <- sample(Homers_data , size = 1000, replace=T)

```

```{r}

library(ggplot2)
sampleNumber = sample(1:10000, size = 1000)
library(dplyr)
my1000 <- Homers_data[sampleNumber, ]
names(my1000)
# Basic density
p <- ggplot(my1000, aes(x=kkardashtemp,col=sex)) + 
  geom_density()
p
# Add mean line
p+ geom_vline(aes(xintercept=mean(kkardashtemp)),
            color="blue", linetype="dashed", size=1)
print(paste('mean = ', mean(my1000$kkardashtemp),' sd = ',sd(my1000$kkardashtemp)))
```

```{r}
numb=2
Hdata = Homers_data[sample(1:1000, size=numb*1000, replace=T),"kkardashtemp"]
mdata  <-matrix(Hdata,nrow=1000,ncol=numb)
x.avg <- apply(mdata,1,mean)
my1000$k2 = x.avg

p <- ggplot(my1000, aes(x=k2)) + 
  geom_density() 
p

# Add mean line
  geom_vline(aes(xintercept=mean(k2)),
            color="blue", linetype="dashed", size=1)
print(paste('mean = ', mean(my1000$k2),' sd = ',sd(my1000$k2)))

```
## We expect the sample mean to be close to population and the standev to get smaller by the ratio of one to the sample size. In this case 1/sqrt(2)

### There are 3 peaks in the density plot. You pick two people you can get MM,MF.FM,or FF. these are all equally likey so the little peak on left is all the FF's while the little peak on the right is all the MM's so the big one in the middle is both the MF and FM in the same bag so-to-speak.  

```{r}
numb = 4
Hdata = Homers_data[sample(1:1000,size=numb*1000, replace=T),"kkardashtemp"]
mdata  <-matrix(Hdata,nrow=1000,ncol=numb)
x.avg <- apply(mdata,1,mean)
my1000$k4 = x.avg


p <- ggplot(my1000, aes(x=k4),col=sex) + 
  geom_density() +

# Add mean line
 geom_vline(aes(xintercept=mean(k4)),
            color="blue", linetype="dashed", size=1)
p + ggtitle("n = 4") +
  xlab("temperature") + ylab(" density")

print(paste('mean = ', mean(my1000$k4),' sd = ',sd(my1000$k4)))

```

### In this case the stdev was cut in half because the sample size was four. This is what the central limit theorem says. There are also 5 peaks. They go from FFFF,to FFFM, to FFMM to FMMM and finally MMMM.

# Now finish by adding blocks with numb being 9, then 16, 25, and 36. be sure to mention how the standard deviation progresses as the sample size changes,


```{r}
numb = 9
Hdata = Homers_data[sample(1:1000,size=numb*1000, replace=T),"kkardashtemp"]
mdata <-matrix(Hdata,nrow=1000,ncol=numb)
x.avg <- apply(mdata,1,mean)
my1000$k3 = x.avg


p <- ggplot(my1000, aes(x=k3),col=sex) +
geom_density() +

# Add mean line
geom_vline(aes(xintercept=mean(k3)),
color="blue", linetype="dashed", size=1)
p + ggtitle("n = 9") +
xlab("temperature") + ylab(" density")

print(paste('mean = ', mean(my1000$k3),' sd = ',sd(my1000$k3)))

```

This graph displays the density distribution of average temperature data collected by Homer. The x-axis represents the average temperature, calculated from 9 data points each, in a sample of 1000 data points. The y-axis shows the density, indicating the frequency of each average temperature. The blue dashed line represents the overall mean of the average temperatures. The standard deviation shrinks as the number gets higher example 4 is divided by 2, 9 is divided by 3. 16 divided by 4 and so on.
The first standard deviation was 41 and now on the 5th one its getting cut down to a third of that total which we got 13.68 but we expect: `r sd(my1000$k3)`.


```{r}
numb = 16
Hdata = Homers_data[sample(1:1000,size=numb*1000, replace=T),"kkardashtemp"]
mdata <-matrix(Hdata,nrow=1000,ncol=numb)
x.avg <- apply(mdata,1,mean)
my1000$k4 = x.avg


p <- ggplot(my1000, aes(x=k4),col=sex) +
geom_density() +

# Add mean line
geom_vline(aes(xintercept=mean(k4)),
color="blue", linetype="dashed", size=1)
p + ggtitle("n = 16") +
xlab("temperature") + ylab(" density")

print(paste('mean = ', mean(my1000$k4),' sd = ',sd(my1000$k4)))
```

As the sample size increases from 9 to 16, the standard deviation of the average temperatures might change. Larger sample sizes tend to result in more precise estimates of the true mean, potentially leading to a decrease in the standard deviation. This code creates a density plot for 1000 rows of temperature data, each calculated from a sample of 16 data points. The standard deviation is now a quarter of what it was which is 10.31 and we expected `r sd(my1000$k4)`.

```{r}
numb = 25
Hdata = Homers_data[sample(1:1000,size=numb*1000, replace=T),"kkardashtemp"]
mdata <-matrix(Hdata,nrow=1000,ncol=numb)
x.avg <- apply(mdata,1,mean)
my1000$k5 = x.avg


p <- ggplot(my1000, aes(x=k5),col=sex) +
geom_density() +

# Add mean line
geom_vline(aes(xintercept=mean(k5)),
color="blue", linetype="dashed", size=1)
p + ggtitle("n = 25") +
xlab("temperature") + ylab(" density")

print(paste('mean = ', mean(my1000$k5),' sd = ',sd(my1000$k5)))
```

As the sample size increases from 16 to 25, the standard deviation of the average temperatures may decrease. Larger sample sizes generally lead to more precise estimates of the true mean, resulting in a narrower distribution around the mean. The density plot with a sample size of 25 might show reduced variability compared to the plot with a sample size of 16, as the larger sample allows for better representation of the underlying population. The standard deviation is now one fifth of what it was in the beginning we got 8 and we expected`r sd(my1000$k5)` and it is only off by a little bit. Our hypothesis was incorrect but not as far along as we thought. 

```{r}
numb = 36
Hdata = Homers_data[sample(1:1000,size=numb*1000, replace=T),"kkardashtemp"]
mdata <-matrix(Hdata,nrow=1000,ncol=numb)
x.avg <- apply(mdata,1,mean)
my1000$k6 = x.avg


p <- ggplot(my1000, aes(x=k6),col=sex) +
geom_density() +

# Add mean line
geom_vline(aes(xintercept=mean(k5)),
color="blue", linetype="dashed", size=1)
p + ggtitle("n = 36") +
xlab("temperature") + ylab(" density")

print(paste('mean = ', mean(my1000$k6),' sd = ',sd(my1000$k6)))
```

As the sample size increases from 16 to 36, the standard deviation of the average temperatures is expected to decrease. This is because larger sample sizes tend to better represent the population, resulting in more precise estimates of the true mean. Therefore, the density plot with a sample size of 36 is likely to have a narrower distribution around the mean, reflecting reduced variability in the data compared to the plot with a sample size of 16. We expect 6 and we got `r sd(my1000$k6)`. My expected value was off by a little, it rounds up to 7 with the code.






























